# Mapping in R

Two weeks ago we covered how to work with data in R and produce some plots.  Today we will learn how to do mapping in R. Let's start by using ggplot2 to plot our arrest data by longitude and latitude. 

## GGPLOT map 

How about we make a map with GGPLOT? We do this by using our longitude and latitude variables as our x and y mapping. 

```{r ggplot-map}
library(ggplot2)
library(tidyverse)
arrests <- read_csv('data/aug6_12_arrest_data.csv')
ggplot(arrests, aes(x=longitude, y=latitude)) + geom_point()
```

We can see the general shape of LA county, but there are limits to what you can do with geospatial data in ggplot. It doesn't know what to do with the latitude and longitude beyond adding the points to a plot. It has no sense of the layers we used in QGIS. But other R packages designed for gis & mapping give us these capabilities. 

## Mapping in R 

We need a way to tell arrests that it is a spatial object in R. To do that we need to use a spatial package called `sf` to do so. Let's load it. 

How do we load packages in R?

```{r}
library(sf)
```


## Converting data to a spatial object

`sf` stands for **simple features** and is based on a standard that specifies how to store and work with 2d geometries, points, lines, polygons, etc.. 

Ok, first up, let's convert our `arrests` data to a simple feature object, so it knows it contains spatial information. We need to do this b/c `tmap` will need this information so we map it. To store this object, I'll create a new data frame called `arrests_sf` and use the `st_as_sf()` that lets us convert our data frame to a `sf` object using the latitude and longitude. Let's look up this funtion in the R help system first. 

```{r help-st_as_sf}
?st_as_sf
```

Many of these conversion functions in the sf package start with `st` - this is because PostGis, a heavily used GIS database uses `ST` (spatial type) for many of its functions. 

Ok, let's now convert our arrests data to a `sf` object. 

```{r}
arrests_sf <- st_as_sf(arrests, coords = c("longitude", "latitude"), crs = 4326)
```

Notice a couple of things you haven't seen before. We have a `coords` argument that we use a `c()` function to **combine** our `latitude` and `longitude` columns. We also tell the function what **Cordinate Refernce System** be are using `4326` is a common one.

Let's look at this new spatial data frame: 

```{r spatial-data-view, eval=FALSE}
View(arrests_sf)
```

We now have a column called `geometry` at the end of our data frame. This contains our latitude and longitude. 

One thing nice is that we can use the `plot()` function that is part of `sf` package to plot our spatial data frame as it is. 

```{r plot-sf}
plot(arrests_sf)
```

But the problem is that we still have no context for our points. We need the layers of polygons we had in QGIS to let us know this is LA. Let's look at this in a package called `tmaps` (thematic maps). 

What do we need to do before we use a package in R? 

```{r load-tmap}
library(tmap)
```

`tmap` let's us plot maps like `ggplot` above -- we'll learn to add different layers to the map, this will be a code way of doing what we did in QGIS. 

```{r}
tm_shape(arrests_sf) + tm_dots()
```

## Reading in shapefiles

Nice, but again, we lack geographic boundaries and other shapes. Just like in QGIS, however, we can read in various shapefiles -- the ones ending in `.shp` -- to provide our map some boundaries. We do this by using the function `st_read`. Let's look at our help documentation on `st_read`. 

Inside our `data/` folder we have the shape files we used in our `QGIS` part of the class plus a new one on the LA county boundaries. Let's read them in individually and note the output in the console, especially the `geometry type`. 

```{r la_county}
#geometry type:  MULTILINESTRING
la_county <- st_read(dsn ="data/DRP_COUNTY_BOUNDARY/DRP_COUNTY_BOUNDARY.shp")
```
```{r zip-coes-streets}
la_zips_st <- st_read(dsn = 'data/CAMS_ZIPCODE_STREET_SPECIFIC/CAMS_ZIPCODE_STREET_SPECIFIC.shp')
```

```{r la_zips}
#geometry type:  MULTIPOLYGON
la_zips <- st_read(dsn = "data/Los_Angeles_City_Zip_Codes/Los_Angeles_City_Zip_Codes.shp")
```

```{r la_freeways}
#geometry type:  MULTILINESTRING
la_freeways <- st_read(dsn ="data/CAMS_FREEWAY_SHIELDS/CAMS_FREEWAY_SHIELDS.shp")
```

Now let's use the layers. The important thing is we match our shapefile with the geometry type. `tm_polygons()` goes with `#geometry type:  MULTIPOLYGON` and `tm_line()` goes with `#geometry type:  MULTILINESTRING`. Let's use this new knowledge to make the map better. 

```{r tmap-mode-plot, echo=FALSE}
tmap_mode("plot")
```

We are going to build this up, let's start with our points like above.

```{r arrests-dots}
tm_shape(arrests_sf) +
  tm_dots() 
```

Now, if the `tmap` function that made our arrests data points on the plot is `tm_dots()`, what do you think the function is for making the zipcode boundaries. What shape are they? 

`tmap` gives a number of functions that lets us determine how we want to map objects. You start with `tm_shape` and that's the function that takes in the data you want to map. Then you tell `tmap` how to map it via functions like `tm_fill`, `tm_polygons`, `tm_borders`, `tm_lines`, `tm_dots` & `tm_bubbles`.  There are many more. Which one you use is dependent on what types of shapes you want to create.  

Let's use our zipcode shapes. 

```{r fill-funtion}
tm_shape(la_zips) +
  tm_fill()
```

Note the output indicates the shape is invalid. This can be for various reasons, but an easy remedy is use the `st_make_valid()` in the lwgeom package. Run the below code. Paste in the etherpad. 

```{r valid-data}
la_zips <- lwgeom::st_make_valid(la_zips)
```
`tm_fill` fills our polygons in zipcodes. Let's see what happens if we use `tm_polygons` instead and add our arrests data as `tm_dots`. Note, we use the plus sign to add layers. 

```{r arrests-la-zips-poly}
tm_shape(arrests_sf) +
  tm_dots() +
tm_shape(la_zips) +
  tm_polygons()
```

Where are our points? Why don't they show? What do you think? 

Yes, they are hidden underneath our zipcodes! WE can reorder the code and they'll appear, but we can also add transparency to our `tm_polygons` function. Does anyone remember what we called this in QGIS? 

Right, `alpha` is the term for adding transparency in plotting and graphing across many software tools, `tmap` is no diff. Let's add a transparency. 

```{r arrests-la-zips-poly-alpha}
tm_shape(arrests_sf) +
  tm_dots() +
tm_shape(la_zips_st) +
  tm_polygons(alpha = 0.5) #a value from 0 to 1
```

Ok, now I want you to add another layer. Add either `la_county` boundary (a polygon) or `la_freeways` to the graph. Take 5min. 

```{r la-zips}
tm_shape(la_zips) +
  tm_polygons() +
tm_shape(la_freeways) +
  tm_lines() +
tm_shape(arrests_sf) +
  tm_dots() 
```

Ok, what we've done so far. 

1. We've taken a dataset that has latitude and longitude data in it and converted it to a simple feature object in R.
2. We've read in shapefiles into R as simple feature objects that become mappable.
3. We've layered them on a plot using `tmap` and even added some transparency to our polygon layer. 

Good job! 

## Aesthetics 

So far, we've accepted `tmap`'s default aesthetic settings, like the grey color fill for `tm_polygons`. However, we can alter this with arguments in our `tm_` function. 

```{r color-polygon}
tm_shape(la_zips) +
  tm_polygons(col='blue')
```

Let's add an alpha:

```{r color-polygon-alpha}
tm_shape(la_zips) +
  tm_polygons(col='blue', alpha = 0.5) +
tm_shape(arrests_sf) +
  tm_bubbles(size = 0.09, col="gold", alpha = 0.4)
```

Instead of `tm_polygons` there's another function that matches our shapes better. It's `tm_borders`. Let's swap it out for `tm_polygons`:

```{r tm-borders-color}
tm_shape(la_zips) +
  tm_fill(col='blue', title="LA Zipcodes", alpha = 0.5) +
    tm_borders()  +
tm_shape(arrests_sf) +
  tm_dots() +
  tm_compass(type = "4star", position = c("left", "top"), size = 2) 
```

## Saving map objects

We can also save a `tmap` as an object in R and reuse it or build on it. For instance:

```{r tmap-save}
lazips <- tm_shape(la_zips) +
    tm_borders()
```

```{r runlazips}
lazips
```

We can now build off of `lazips` and then use different versions of the map. 

```{r adding-tolazips}
lazips <- tm_shape(la_zips) 
lzip1 <- lazips + tm_fill(col = "red")
lzip2 <- lazips + tm_fill(col = "red", alpha = 0.3)
lzip3 <- lazips + tm_borders(col = "blue")
lzip4 <- lazips + tm_borders(lwd = 3)
```

```{r print-out-base-no-layer}
lazips #our familiar map
```


```{r print-out-map}
lzip1
lzip2
lzip3
```

```{r arrange-maps}
tmap_arrange(lzip1, lzip2, lzip3, lzip3)
```

`tmap` is very flexible and granular. We can control all aspects of the map and as we move along we can show more 

```{r layout}
lzip1 + tm_layout(title="LA Zipcodes")
```

```{r la-zip-style}
lzip3 + tm_style("cobalt") + tm_layout(title="LA Zips", title.color = "white")
#tm_style("classic")
#tm_style("bw")
#tm_style("col_blind")
#look at ?tm_layout for more
```

## Interactive Maps

So far, we've been using `tmap` to produce a static map, but `tmap` lets us create interactive maps by switching it's mode to `view'. We can do this using a single line. 

```{r tmap-mode}
tmap_mode("view")
```

```{r tmap-view}
tm <- tm_shape(la_zips) +
  tm_polygons() +
tm_shape(la_freeways) +
  tm_lines() +
tm_shape(arrests_sf) +
  tm_dots() 
tmap_save(tm, filename = "world_map.html")
```
[aggre-larger-data]:# use dplyr to aggregate data by zipcode, then create a spatial dataset, by sex, by race, by age
[map data]:# look at SpaData package on how to set up spatial data as a package - add variants of arrests data
[add-color-other-theme]:# add color to the plot and different data, will need to prep
[write challenges]:# focus on faded examples - fill in the blank style
[write learning objectives]:# what do we want them to get out of the next few classes, how to make a map in tmap, how to convert data to publish data, how to add color elements
[tmap-functions]:# add color via tm_border, tm_fill https://geocompr.robinlovelace.net/adv-map.html#aesthetics 
[tmap-functions]:#

