# Mapping in R: tmaps continued

Two weeks ago we covered some basic 

## GGPLOT map 

How about we make a map with GGPLOT? We do this by using our longitude and latitude variables as our x and y mapping. 

```{r}
ggplot(arrests, aes(x=longitude, y=latitude)) + geom_point()
```

We can see the general shape of LA county, but there are limits to what you can do with geospatial data in ggplot. It doesn't know what to do with the latitude and longitude beyond adding the points to a plot. It has no sense of the layers we used in QGIS. But other R packages gives us these capabilities. 

## Mapping in R 

We need a way to tell arrests that it is a spatial object in R. To do that we need to use two spatial packages. Let's load them. 

```{r}
library(sf)
library(tmap)
```

`sf` stands for spatial features. `tmap` let's us plot maps like `ggplot`. Ok, first up, let's convert our `arrests` data to a spatial feature. We need to do this b/c `tmap` will need this information so we can use on a map. I'll create a new data frame called `arrests_sf` and use the `st_as_sf()` that lets us convert our data frame to a `sf` object using the latitude and longitude. 

```{r}
arrests_sf <- st_as_sf(arrests, coords = c("longitude", "latitude"), crs = 4326)
```

Notice a couple of things you haven't seen before. We have a `coords` argument that we use a `c()` function to **combine** our `latitude` and `longitude` columns. We also tell the function what **Cordinate Refernce System** be are using `4326` is a common one. 

Let's look at this new spatial data frame: 

```{r, eval=FALSE}
View(arrests_sf)
```

We now have a column called `geometry` at the end of our data frame. This contains our latitude and longitude. 

One thing nice is that we can use the `plot()` function to plot our spatial data frame. 

```{r}
plot(arrests_sf)
```

But the problem is that we still have no context for our points. We need the layers of polygons we had in QGIS to let us know this is LA. Let's look at this in a package called `tmaps` (thematic maps). 

```{r}
tm_shape(arrests_sf) + tm_dots()
```

Nice, but again, we lack geographic boundaries and other shapes. Just like in QGIS, however, we can read in various shapefiles -- the ones ending in `.shp` -- to provide our map some boundaries. We do this by using the function `st_read`. Let's look at our help documentation on `st_read`. 

Inside our `data/` folder we have the shape files we used in our `QGIS` part of the class plus an new one on the LA county boundaries. Let's read them in individually and note the output, especially the `geometry type`. 

```{r la_county}
#geometry type:  MULTILINESTRING
la_county <- st_read(dsn ="data/DRP_COUNTY_BOUNDARY/DRP_COUNTY_BOUNDARY.shp")
```

```{r la_zips}
#geometry type:  MULTIPOLYGON
la_zips <- st_read(dsn = "data/Los_Angeles_City_Zip_Codes/Los_Angeles_City_Zip_Codes.shp")
```

```{r la_freeways}
#geometry type:  MULTILINESTRING
la_freeways <- st_read(dsn ="data/CAMS_FREEWAY_SHIELDS/CAMS_FREEWAY_SHIELDS.shp")
```

Now let's use the layers. The important thing is we match our shapefile with the geometry type. `tm_polygons()` goes with `#geometry type:  MULTIPOLYGON` and `tm_line()` goes with `#geometry type:  MULTILINESTRING`. Let's use this new knowledge to make the map better. 

```{r tmap-mode-plot, echo=FALSE}
tmap_mode("plot")
```



```{r}
tm_shape(la_zips) +
  tm_polygons() +
tm_shape(la_freeways) +
  tm_lines() +
tm_shape(arrests_sf) +
  tm_dots() 
```

Ok, that's more like it. Last thing before we break. So far, we've been using `tmap` to produce a static map, but `tmap` lets us create interactive maps by switching it's mode to `view'. We can do this using a single line. 

```{r tmap-mode, echo=FALSE}
tmap_mode("view")
```

```{r tmap-view, eval=FALSE}
tm_shape(la_zips) +
  tm_polygons() +
tm_shape(la_freeways) +
  tm_lines() +
tm_shape(arrests_sf) +
  tm_dots() 
```
[]:# - come up with R script has fill in the blanks challenges 
[]:#maybe an example with tmap 
[]:# assign homework for the primers in dataviz few, boxplot, a couple more
