# Mapping with R Continued

```{r load-libs}
library(tidyverse)
library(sf)
library(tmap)
```


```{r read-setup-data-shp-files, message=FALSE}
arrests <- read_csv('data/aug6_12_arrest_data.csv')
arrests_sf <- st_as_sf(arrests, coords = c("longitude", "latitude"), crs = 4326)
#geometry type:  MULTILINESTRING
la_county <- st_read(dsn ="data/DRP_COUNTY_BOUNDARY/DRP_COUNTY_BOUNDARY.shp")
la_zips_st <- st_read(dsn = 'data/CAMS_ZIPCODE_STREET_SPECIFIC/CAMS_ZIPCODE_STREET_SPECIFIC.shp')
#geometry type:  MULTIPOLYGON
la_zips <- st_read(dsn = "data/Los_Angeles_City_Zip_Codes/Los_Angeles_City_Zip_Codes.shp")
#geometry type:  MULTILINESTRING
la_freeways <- st_read(dsn ="data/CAMS_FREEWAY_SHIELDS/CAMS_FREEWAY_SHIELDS.shp")
## the below command will make our zip data valid
la_zips <- lwgeom::st_make_valid(la_zips)
```

## Revisit last week 

Last week we finished with this map.

```{r tmap-mode-change-view-plot, echo=FALSE}
tmap_mode("plot")
```

```{r arrests-bubbles-compass}
 tm_shape(la_zips) +
   tm_polygons(col="blue", alpha = 0.3) +
   tm_shape(arrests_sf) +
   tm_bubbles(size = 0.09, col="gold", alpha=0.4)+
   tm_compass(type = "4star", position = c("left", "top"), size = 2) 
```

```{r arrests-bubbles-compass-var-map}
 tm_shape(la_zips) +
   tm_polygons(col="blue", alpha = 0.3) +
   tm_shape(arrests_sf) +
   tm_bubbles(size = 0.09, col="sex", alpha=0.4)+
   tm_compass(type = "4star", position = c("left", "top"), size = 2) 
```

Let's affect the color of the dots by setting a palette argument.

```{r arrests-bubbles-compass-palette}
#using a palette
 tm_shape(la_zips) +
   tm_polygons(col="blue", alpha = 0.3) +
   tm_shape(arrests_sf) +
   tm_bubbles(size = 0.09, col="sex", palette = "Set3")+
   tm_compass(type = "4star", position = c("left", "top"), size = 2) 
```

We can find the available palettes via this code: 

```{r palette-explore, eval=FALSE}
#lets look at options
tmaptools::palette_explorer()
```

Try out a few palettes. 

We can also add a style argument.  

```{r add-style}
tm_shape(la_zips) +
   tm_polygons(col="blue", alpha = 0.3) +
   tm_shape(arrests_sf) +
   tm_bubbles(size = 0.09, col="sex", style = "cat")+
   tm_compass(type = "4star", position = c("left", "top"), size = 1) 
```


## Saving map objects

We can also save a `tmap` as an object in R and reuse it or build on it. For instance:

```{r tmap-save-object}
lazips <- tm_shape(la_zips) 
```

If we run this we get an error that no layer is define - we haven't told `tmap` how to display our la_zips data. 

```{r run-lazips, eval=FALSE}
lazips
```

**BUT** we can now build off of `lazips` and then use different versions of the map. 

```{r adding-to-lazips-red}
lazips + tm_fill(col = "red")
```

We can use this idea and save our work after each change. Let's create multiple objects with different colors and other parameters. 

```{r adding-tolazips-multi-objs}
lzip1 <- lazips + tm_fill(col = "red")
lzip2 <- lazips + tm_fill(col = "red", alpha = 0.3)
lzip3 <- lazips + tm_borders(col = "blue")
lzip4 <- lazips + tm_borders(lwd = 3)
```


```{r print-out-map-one-by-one}
lzip1
lzip2
lzip3
lzip4
```

Even more interesting we can compose these maps into a grid by using `tmap_arrange` function: 

```{r arrange-maps-lazips}
tmap_arrange(lzip1, lzip2, lzip3, lzip4)
```

The most commonly used aesthetics for fill and border layers include color, transparency, line width and line type, set with `col`, `alpha`, `lwd`, and `lty` arguments, respectively.

`tmap` is very flexible and granular. We can control all aspects of the map and as we move along we can show more 

```{r layout-zips}
lzip1 + tm_layout(title="LA Zipcodes")
```

```{r la-zip-tmstyle}
lzip3 + tm_style("cobalt") + tm_layout(title="LA Zips", title.color = "white")
#tm_style("classic")
#tm_style("bw")
#tm_style("col_blind")
#look at ?tm_layout for more
```

## Faceted Maps (small multiples)

One way to present different data on a map is to split the data by a category and create two maps with different data. Let's look at this example. 

```{r faceted-map}
tm_shape(la_zips) +
   tm_polygons() +
   tm_shape(arrests_sf) +
   tm_symbols(col = "black", border.col = "white", size = 0.5) +
   tm_facets(by = "sex", nrow=2, free.coords = FALSE)
```

Let's color by multiple variables in our data: 

```{r faceted-map-two-var}
tm_shape(la_zips) +
   tm_polygons() +
   tm_shape(arrests_sf) +
   tm_symbols(col = "race_cat", border.col = "white", size = 0.5) +
   tm_facets(by = "sex", nrow=2, free.coords = FALSE)
```


## Interactive Maps

So far, we've been using `tmap` to produce a static map, but `tmap` lets us create interactive maps by switching it's mode to `view` instead of the default `plot`. We can do this using a single line. 

```{r tmap-mode-change-view}
tmap_mode("view")
```

```{r tmap-view-zip-freeway-layers}
tm <- tm_shape(la_zips) +
  tm_polygons() +
tm_shape(la_freeways) +
  tm_lines() +
tm_shape(arrests_sf) +
  tm_dots() 
#tmap_save(tm, filename = "world_map.html")
```

```{r eval=FALSE}
tm
```

Note, that we can interact with the map and change the base layer, turn off and on the layers. 

## Publishing the map using rpubs 

Open a RMarkdown document, copy 

[aggre-larger-data]:# use dplyr to aggregate data by zipcode, then create a spatial dataset, by sex, by race, by age
[add-color-other-theme]:# add color to the plot and different data, will need to prep
[tmap-functions]:# add color via tm_border, tm_fill https://geocompr.robinlovelace.net/adv-map.html#aesthetics 
[shinyjs]:# need to install shinyjs to run palette explorer

