---
title: 'AF AM 188: R Assignment 2'
author: "Tim Dennis"
date: "11/26/2019"
output: html_document
---

```{r setup-assignment, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

# R Assignment 2  

Please add your name and email to the below lines: 

Name:   
Email:   

## Instructions 

Complete the R coding challenges and questions in RStudio Cloud Project
<https://rstudio.cloud/spaces/31925/project/757350> in the `r-assignment-2.R` file.
Save that file as you work. You will have until **12/06 by 5pm**. If you encounter any problems, please come to
office hours on 12/3 from 2-4pm in the Data Science Center, 21536 YRL, email us at
<datascience@ucla.edu> or set up an [appointment](https://calendly.com/data-science-team/dsc-consulting?back=1&month=2019-11). 

**This assignment is worth 32 points.**

## 1. For this exercise we need to load packages in R.  (3 points)

```{r 1-load-libraries}
#load libraries we need. Also, set tmap_mode to plot b/c we want static maps for this exercise.
library(tidyverse)
library(sf)
library(tmap)
tmap_mode("plot")
```


## 2.  We also will need to load the data we will be using to map in R. 
Load the data. (3 points)

```{r 2-load-data, message=FALSE}
arrests <- read_csv('data/aug6_12_arrest_data.csv')
arrests_sf <- st_as_sf(arrests, coords = c("longitude", "latitude"), crs = 4326)
#geometry type:  MULTILINESTRING
la_county <- st_read(dsn ="data/DRP_COUNTY_BOUNDARY/DRP_COUNTY_BOUNDARY.shp")
la_zips <- st_read(dsn = "data/Los_Angeles_City_Zip_Codes/Los_Angeles_City_Zip_Codes.shp")
#geometry type:  MULTILINESTRING
## the below command will make our zip data valid
la_zips <- lwgeom::st_make_valid(la_zips)
la_freeways <- st_read(dsn ="data/CAMS_FREEWAY_SHIELDS/CAMS_FREEWAY_SHIELDS.shp")
```

## 3. Altering a bubble map. (6pts)

In class we made the below map using `tmap`.  

```{r 3-bubbles-map}
tm_shape(la_zips) +
   tm_polygons(col="blue", alpha = 0.3) +
   tm_shape(arrests_sf) +
   tm_bubbles(size = 0.09, col="gold", alpha=0.4)+
   tm_compass(type = "4star", position = c("left", "top"), size = 2) 
```

Alter the below code by filling in the blanks with a different color for the
`tm_polygons` and `tm_bubbles` functions. Also, adjust the `alpha` parameters. What
colors are available for you to use? You can run the below function `colors()` for a
list: 

```{r see colors, eval=FALSE}
colors()
```


```{r 3-alter-bubble-map, eval=FALSE}
tm_shape(la_zips) +
   tm_polygons(col="___", alpha = ___) +
   tm_shape(arrests_sf) +
   tm_bubbles(size = 0.09, col="___", alpha= ___)+
   tm_compass(type = "4star", position = c("left", "top"), size = 2) 
```


## 4. Mapping arrests by sex (6pts)

Changing the map above so the bubbles are colored based on the sex variable in our
data. Fill in the blank in the `tm_bubbles` function. 

```{r arrests-by-sex, eval=FALSE}
tm_shape(la_zips) +
   tm_polygons(col="blue", alpha = 0.3) +
   tm_shape(arrests_sf) +
   tm_bubbles(size = 0.09, col="___", palette=c(F='cyan', M='red')) +
   tm_compass(type = "4star", position = c("left", "top"), size = 2) 
```

What does the map tell you about the data? 

Answer: 

## 5. Using arrange to make a row of maps (7 points)

With the below code, we create multiple maps based off the base map `lazips`. 
Run the code to create the R objects. 

```{r creating-maps-base-map}
lazips <- tm_shape(la_zips)
lzip1 <- lazips + tm_fill(col = "red")
lzip2 <- lazips + tm_fill(col = "red", alpha = 0.3)
lzip3 <- lazips + tm_borders(col = "blue")
lzip4 <- lazips + tm_borders(lwd = 3)
```

Now, use the function `tmap_arrange()` to arrange `lzip1` to `lzip4`, printing them out in a row of maps in the right Plots pane. To see how we did this in class, see section 5.2 of the lesson: https://afam188.netlify.com/mapping-with-r-continued.html#saving-map-objects-1

```{r using-arrange-maps, eval=FALSE}
tmap_arrange(lzip1, ____, _____, ____, ____)
```


## 6. Creating a faceted map (7 points)

In class we produced a faceted map of arrests by sex with the below code. 
Run the code to see how it works: 

```{r faceted-by-sex}
tm_shape(la_zips) +
   tm_polygons() +
   tm_shape(arrests_sf) +
   tm_symbols(col = "black", border.col = "white", size = 0.5) +
   tm_facets(by = "sex", nrow=2, free.coords = FALSE)
```

Now, for below, fill in the blank in the `tm_facets` function to produce a faceted map
of arrests by race category `race_cat`. This should produce a plot for each racial
category with arrest points colored by race. 

```{r arrests-by-facets, eval=FALSE}
tm_shape(la_zips) +
   tm_polygons() +
   tm_shape(arrests_sf) +
   tm_symbols(col = "sex", border.col = "white", size = 0.5, palette=c(M = 'yellow', F = "steelblue")) +
   tm_facets(by = "____", nrow=2, free.coords = FALSE)
```
